---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
The data is loaded from the file contained in the .zip file, using the *unzip* then the *read.csv* functions. 

Then the data is processed as total steps per day, as well as mean and median of steps per day, using the *aggregate* function.

```{r}
unzip("activity.zip")
data <- read.csv("activity.csv")

perDay <- aggregate(data[[1]], by=data[2], sum)
##perDay <- cbind(perDay, mean=aggregate(data[[1]], by=data[2], mean)[[2]], median=aggregate(data[[1]], by=data[2], median)[[2]])
name()
```

## What is mean total number of steps taken per day?

An histogram is shown for the total number of steps taken per day.

```{r}
hist(perDay[[2]], xlab= "Total number of steps per day")
```

The mean and median total number of steps per day may be shown using the *summary* function

```{r}
summary(perDay[[2]])
```

## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
